---
title: "parallel"
author: "Kun Yang & Alexander Zhu"
date: "3/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# source the rlnp function
source("../R/rlnp.R")
# check the length of rlnpx to see it should return a vector of 1000 draws 
length(rlnpx(1000, 10, 0.5))
```

`First, we investigate the elapsed time for computing rlnpx(5e5,10,1). It is noted here that we don't compute rlnpx(5e6,10,1) as asked given the long running time in local server. 
```{r}
# system.time(rlnpx(5e6, 10, 1))
# Error: cannot allocate vector of size 1.9 Gb
# try 5e5 instead
n <- 5e5
system.time(rlnpx(n, 10, 1))
```

```{r}
n5 <- n/5
system.time(unlist(lapply(list(n5,n5,n5,n5,n5), rlnpx, y=10, scale=1)))
```
Then we compare it with the use of parallel library. 
```{r}
library(parallel)
system.time(unlist(mclapply(list(c1=n5,c2=n5,c3=n5,c4=n5,c5=n5), rlnpx, y=10, scale=1, mc.cores=3L)))
```
The result suggests the elapsed time reduced significantly from 11.585 to 6.1. This confirms the power of using parallel library to reduce the elapsed time. 


## Contribution Statement

Both Alex and Kun wrote their own code of the function first and discuss on the final version.

Repository: [kuy003](https://github.com/Kun-Yang-003/FMPH291_WI22_A00_kuy003.git)
