---
title: "Introduction to Package 'pareto'"
author: "Kun Yang & Alexander Zhu"
date: "2/17/2022"
output: pdf_document
---

The Pareto distribution density is calculated using the following formula: $$ f(x|\alpha,\beta) = \frac{\beta{\alpha}^\beta}{x^{\beta+1}} $$
for $0<\alpha\leq x$ and $\beta>0$. Note that densities are defined on the entire real line; they are zero outside of the support of the distribution.

## Code
```{r, eval=FALSE}
Rcpp::sourceCpp("../Rcpp/dpareto.cpp")
source("../R/dpareto.R")

ppareto <- function(q, alpha, beta, lower.tail = TRUE, log.p = FALSE){
  max_length <- max(length(q), length(alpha), length(beta))
  alpha <- rep(alpha, length = max_length)
  beta <- rep(beta, length = max_length)
  q <- rep(q, length = max_length)
  log_p <- suppressWarnings(ifelse(q<alpha, 0, beta*(log(alpha)-log(q))))
  log_p <- suppressWarnings(ifelse(alpha <=0 | beta <=0, NaN, log_p))
  if (any( is.nan( log_p )) == TRUE){
    warning("NaNs produced")
  }
  p <- log_p
  if(! log.p) p <- exp(log_p)
  if(lower.tail) p <- 1 - p
  if(log.p & (! lower.tail)) p <- log(1-exp(log_p))
  p
}

qpareto <- function(p, alpha, beta, lower.tail = TRUE, log.p = FALSE){
  max_length <- max(length(p), length(alpha), length(beta))
  alpha <- rep(alpha, length = max_length)
  beta <- rep(beta, length = max_length)
  p <- rep(p, length = max_length)
  if(log.p) p = exp(p)
  if(lower.tail) p = 1 - p
  q <- suppressWarnings(alpha/(p^(1/beta)))
  q <- suppressWarnings(ifelse(alpha <=0 | beta <=0, NaN, q))
  if (any( is.nan( q )) == TRUE){
    warning("NaNs produced")
  }
  q
}
```

```{r}
library(tidyverse)
## R CMD INSTALL pareto_1.1.tar.gz
library(pareto)
```

## Usage
```{r}
dpareto(x = 2:4, alpha = 1, beta = 2)
ppareto(q = 2:4, alpha = 1, beta = 2)
qpareto(p = c(0.1, 0.5, 0.9), alpha = 1, beta = 2)
```

`dpareto` gives the density,  
`ppareto` gives the distribution function,  
`qpareto` gives the quantile function.  



```{r}
df_p <- data.frame(q = seq(1, 10, 0.1), alpha = 1, beta = 1)
df_p <- df_p %>% mutate(beta = 2) %>% rbind(df_p, .)
df_p <- df_p %>% mutate(alpha = 2) %>% rbind(df_p, .)
df_p <- df_p %>% mutate(den = dpareto(q, alpha, beta), cum = ppareto(q, alpha, beta)) %>% mutate(alpha = as.factor(alpha), beta = as.factor(beta))
```

## Density
```{r}
ggplot(df_p, aes(x = q, y = den, color = alpha, linetype = beta)) + 
  geom_line() + 
  theme_bw()
```

## Cumulative Distribution
```{r}
ggplot(df_p, aes(x = q, y = cum, color = alpha, linetype = beta)) + 
  geom_line() + 
  theme_bw()
```


```{r eval=FALSE, include=FALSE}
## for creating R package
working.Rdir <- file.path("packages", "working", "R")
working.Rfiles <- file.path(working.Rdir, c("dpareto.R", "ppareto.R", "qpareto.R"))
package.path <- file.path("packages")
# package.skeleton("pareto", code_files = working.Rfiles, path = package.path)
Rcpp::compileAttributes(pkgdir = "packages/pareto/",verbose = TRUE)
# Rcpp::compileAttributes
## R CMD build pareto
## R CMD check pareto_1.0.tar.gz
## R CMD INSTALL pareto_1.0.tar.gz
```

## Test code
```{r}
stopifnot(all.equal(dpareto(3,2,1), 0.2222222222))
stopifnot(all.equal(dpareto(1,2,3), 0.0))
stopifnot(all.equal(dpareto(3:5,2, 1), c(0.2222222222, 0.1250000, 0.0800000)))
stopifnot(all.equal(dpareto(6,2:4, 1), c(0.05555555556, 0.08333333333, 0.11111111111)))
stopifnot(all.equal(log(dpareto(1:5,2, 1)), dpareto(1:5,2, 1, log = TRUE)))
stopifnot(all.equal(dpareto(6,1,2:4), c(0.0092592593, 0.0023148148, 0.0005144033)))
stopifnot(all.equal(dpareto(1, 2, 1:2), c(0, 0)))


stopifnot(all.equal(ppareto(3:5,2, 1),c(0.3333333, 0.5000000, 0.6000000), tolerance = 1e-6))
stopifnot(all.equal(ppareto(1,2, 1),0))
stopifnot(all.equal(ppareto(0,0, 1),NaN))

stopifnot(all.equal(qpareto(1,1,1),Inf))
stopifnot(all.equal(qpareto(0.5,1,2), 1.414214, tolerance = 1e-6))
```

## Contribution Statement

Both Alex and Kun wrote their own code of the function first and discuss on the final version.

Repository: [kuy003](https://github.com/Kun-Yang-003/FMPH291_WI22_A00_kuy003.git)

