---
title: "Homework 8"
author: "Kun Yang & Alexander Zhu"
date: "2/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A brief explanation of how rejection sampling works:
The rejection sampling method generates sampling values from a target distribution $X$ with arbitrary probability density function $f(x)$ by using a proposal distribution $Y$ with probability density $g(x)$.  The idea is that one can generate a sample value from $X$ by instead sampling from $Y$ and accepting the sample $Y$ with probability $f(x)/(Mg(x))$, repeating the draws from $Y$ until a value is accepted. $M$ here is a constant and the finite bound on the likelihood ratio $f(x)/g(x)$. The  validation of this method is the envelope principle: when simulating the pair $(x,v=u*Mg(x))$, one produces a uniform simulation over the subgraph of $Mg(x)$. Accepting only pairs such that $u<f(x)/(Mg(x))$ then produces pairs $(x,y)$ uniformly distributed over the subgraph of  and thus, a simulation from $f(x)$.

In this homework,We know the density of proposal distribution(here it is Cauchy distribution) to sample from, and create an envelop which completely encloses the proposed distribution. 

The target function here is the dlnp function which returns the density of $\phi(x) exp(-exp(x))exp(xy)$. 

The envelop here is the random deviates function rlnp which returns samples from a target distribution whose density is proportional to $\phi(x) exp(-exp(x))exp(xy)$ (i.e. the R function dlnp) using a rejection sampler. In detail, the rlnp function should find a curve proportional to a Cauchy density (with scale as its scale parameter) that matches dlnp as nearly as possible and always is at least as high as dlnp. That curve is the envelope. 

## log-Normal Poisson distribution

The density of LNP distribution:  
$$
f(x;y)=\phi(x)\exp(-\exp(x))\exp(xy)
$$

```{r}
dlnp <-
function(x,y ,log = FALSE) {
val <-
dnorm( x, log = TRUE ) +
dpois( y, exp( x ), log = TRUE)
if (log) val else exp( val )
}
```

The mode of the density can be found roughly with `locator` function.  
```{r}
# curve(dlnp(x,2),-5,5)
# grid()
# locator(1)
# c(0.4092479, 0.09262501)
```

```{r, eval=FALSE}
curve(dcauchy(x,0.4092479,1),-5,5)
# 0.3187746
```

### envelope
Envelope is a scaled Cauchy density that approximates a LNP density by timing the relative height of a LNP density curve to a Cauchy density curve.  

```{r}
curve(dlnp(x,2),-5,5)
curve(0.09262501/0.3187746*dcauchy(x,0.4092479,1),-5,5, add = T, col="red")
```



## rlnp & rlnpPlot
```{r}
rlnp <- function(n, y, scale = 1){
  # x <- seq(-5,5,0.001)
  # imax <- which.max(dlnp(x,y))
  # x0 <- x[imax]
  # M <- dlnp(x0,y)/max(dcauchy(x, x0, scale))
  x0 <- optimize(function(x){dlnp(x, y)}, c(-5, 5), maximum = T)$maximum
  M <- dlnp(x0,y)/optimize(function(x){dcauchy(x, x0, scale)}, c(-5, 5), maximum = T)$objective
  x <- c()
  yunif <- c()
  dratio <- c()
  keep <- c()
  n_ <- 10*n
  while(sum(keep) < n){
    x_ <- rcauchy(n_)
    yunif_ <- runif(n_)
    dratio_ <- dlnp(x_, y) / (M*dcauchy(x_, x0, scale))
    keep_ <- yunif_ <= dratio_
    x <- c(x, x_)
    yunif <- c(yunif, yunif_)
    dratio <- c(dratio, dratio_)
    keep <- c(keep, keep_)
  }
  i.end <- max(which(keep)[1:n])
  return(list(x = x[1:i.end], yunif = yunif[1:i.end], dratio = dratio[1:i.end], keep = keep[1:i.end], x0 = x0, M = M, y = y, scale = scale))
}

rlnpPlot <- function(obj){
  with(obj, {
    curve( M * dcauchy( x , x0, scale), min(x[ keep ]), max(x[ keep ], lty = 4), ylab="", xlab = "")
    curve( dlnp( x , y), min(x[ keep ]), max(x[ keep ]), col = 2, add = TRUE, ylab="", xlab = "")
    points( x[ keep ], yunif[ keep ] * (dcauchy( x[ keep ] , x0, scale) * M), pch=".", col=2, ylab="", xlab = "")
    points( x[ ! keep ], yunif[ ! keep ] * (dcauchy( x[ ! keep ] , x0, scale) * M), pch=".", ylab="", xlab = "")
    legend( "topright", col = 1:2, lty = rep( 1, 2 ), legend = c( "denvelope/rejected", "dlnp/accepted"))
    title(main = paste0("dlnp(x, ", y, ")"), sub = paste0("denvelope(x) = ", round(M,2), " * dcauchy(x,", round(x0,2), ",", round(scale,2), ")", "\n", "fraction of samples rejected = ", round(1-sum(keep)/length(keep),2)))
  })
}
```

When the `y` parameter is 0, the envelope doesn't cover the density of LNP well (over 5000 `dratio` values exceeds 1.0), so there is a serious flaw in this sampler.  
```{r}
lst.lnp <- rlnp(10000, 0)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1 <- function(obj, nmax=20){
  dratio_ex <- obj$dratio[which(lst.lnp$dratio>1)]
if(length(dratio_ex) == 0){
  cat("No dratio exceeds 1.0")
}else if(length(dratio_ex) > nmax){
  cat(length(dratio_ex), "dratio values exceeds 1.0")
}else{
  cat(length(dratio_ex), "dratio values exceeds 1.0:", paste(dratio_ex, sep = ", "))
}
}
```

```{r}
dratio_ex1(lst.lnp)
```


When the `y` parameter is 3, the envelope covers the density of LNP from the plot but the fraction of rejection is higher (`M` is smaller).  
```{r}
lst.lnp <- rlnp(10000, 3)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```


When the `y` parameter is 10, the fraction of rejection gets higher.  

```{r}
lst.lnp <- rlnp(10000, 10)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```


For the `y` parameter of 100, almost all generated samples are rejected.  

```{r}
lst.lnp <- rlnp(10000, 100)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```


If we change the scale parameter to 2, more samples get rejected.  

```{r}
lst.lnp <- rlnp(10000, 100, 2)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```


When the scale parameter is 0.1, the envelope doesn't cover `dlnp` completely.  

```{r}
lst.lnp <- rlnp(10000, 100, 0.1)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```


When the scale parameter is 0.2, the envelope seems to cover `dlnp` completely but the rejection fraction is almost 1.  

```{r}
lst.lnp <- rlnp(10000, 100, 0.2)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```


The rejection fraction is smaller when scale parameter equals to 0.15. So we take 0.15 as the `best.scale` for `y=100`. Considering few `dratio` values exceeds 1 in this sampler, 0.2 should be keep as a more conservative choice.  

```{r}
lst.lnp <- rlnp(10000, 100, 0.15)
rlnpPlot(lst.lnp)
```

```{r}
dratio_ex1(lst.lnp)
```



## Contribution Statement

Both Alex and Kun wrote their own code of the function first and discuss on the final version.

Repository: [kuy003](https://github.com/Kun-Yang-003/FMPH291_WI22_A00_kuy003.git)

