---
title: "sparse"
author: "Alexander Zhu & Kun Yang"
date: "1/31/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following instructions are applied for calculating the log-likelihood of a vector $Y$, which has a multivariate normal distribution with mean zero and covariance matrix  
$$
C = 
\begin{bmatrix}
1 & a & 0 & \cdots & 0 & 0 & 0\\ 
a & 1 & a & \cdots & 0 & 0 & 0\\ 
0 & a & 1 & \cdots & 0 & 0 & 0\\ 
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots & \vdots\\ 
0 & 0 & 0 & \cdots & 1 & a & 0\\ 
0 & 0 & 0 & \cdots & a & 1 & a\\ 
0 & 0 & 0 & \cdots & 0 & a & 1
\end{bmatrix}
$$

The log-likelihood for an observed vector $y$, dropping additive constants, is  
$$
-(1/2)\mathrm{log}\,\mathrm{det}C-(1/2)y^TC^{-1}y
$$

## (a) log-likelihood with Cholesky decomposition

First, let's define a function to generate the covariance matrix $C$,  

```{r}
gen_C_1 <- function(n, a){
  C <- diag(nrow = n, ncol = n)
  C[cbind(1:(n-1),2:n)] <- a
  C[cbind(2:n,1:(n-1))] <- a
  C
}
```

_The size parameter `n` and a constant `a` is required._  


By introducing Cholesky decomposition to $C$, we have  

$$
\begin{aligned}
& C = R^TR \\
& -(1/2)\mathrm{log}\,\mathrm{det}C = -\sum\mathrm{log}\,\mathrm{diag}L \\
& y^TC^{-1}y = y^TR^{-1}R^{-T}y = z^Tz \\
& \textrm{where, }R^Tz=y
\end{aligned}
$$

So the function to calculate the log-likelihood of vector $y$ with known $a$ can be defined as follow:  

```{r}
loglikhd_1 <- function(y, a){
  C <- gen_C_1(length(y), a)
  R <- chol(C)
  -sum(log(diag(R)))-0.5*t(forwardsolve(t(R), y)) %*% forwardsolve(t(R), y)
}
```

_The log-likelihood can be calculated by inputting a vector `y` and the known `a`._  


## (b) Using `bandSparse` function from package  

The only difference is the way we create the $C$ matrix.  

```{r}
library(Matrix)
gen_C_2 <- function(n, a) bandSparse(n = n, k = -c(0:1), diag = list(rep(1, n), rep(a, n-1)), symm = T)

loglikhd_2 <- function(y, a){
  C <- gen_C_2(length(y), a)
  R <- chol(C)
  -sum(log(diag(R)))-0.5*t(forwardsolve(t(R), y)) %*% forwardsolve(t(R), y)
}
```


## (c) Simulation  

```{r}
library(MASS)
```

```{r}
n <- 1000
a <- 0.5
C <- diag(nrow = n, ncol = n)
C[cbind(1:(n-1),2:n)] <- a
C[cbind(2:n,1:(n-1))] <- a

set.seed(111)
y <- mvrnorm(1, mu = rep(0, n), Sigma = C)
```


Using original formula:  

```{r}
- 0.5 * log(det(C)) - 0.5 * t(y) %*% solve(C) %*% y
system.time(- 0.5 * log(det(C)) - 0.5 * t(y) %*% solve(C) %*% y)
```

Using the method in part (a):  

```{r}
loglikhd_1(y, a)
system.time(loglikhd_1(y, a))
```

Using the method in part (b):  

```{r}
loglikhd_2(y, a)
system.time(loglikhd_2(y, a))
```

All 3 methods return same result, and the last method is most efficient.  

# Contribution Statement
Both Alex and Kun wrote their own code of the function first and discuss on the final version.

Repository: [kuy003](https://github.com/Kun-Yang-003/FMPH291_WI22_A00_kuy003.git)

