---
title: "HW6"
author: "Kun Yang & Alexander Zhu"
date: "2/14/2022"
output: pdf_document
---

To use EM algorithm, we first write an R function called tabll that will compute the log likelihood: 
```{r}
tabll <- function(p.vec, eta.mat, X){
  # X is a N by C matrix of counts
  N <- nrow(X)
  C <- ncol(X)
  # p.vec is a vector of length M containing (p1,p2,...,pm)
  M <- length(p.vec)
  # check condition that eta.mat is a M by C matrix 
  if(ncol(eta.mat) != C) warning("wrong dimension")
  if(nrow(C) != M) warning("wrong dimension")
  m.vec <- rowSums(X)
  # log likelihood function 
  ll <- 1
  for(i in 1:N){
  lli <- 0
  for(j in 1:M){
    lli <- lli + p.vec[j] * log(dmultinom(X[i,], prob = eta.mat[j,]))
  }
  ll <- ll * lli
  }
  ll
}
```

Then we move to step 4. We write a function ExZ.X, to compute the expected value of Z. We start with a function prop.by.row to show the proportion of each row. 

```{r}
# prop.by.row <- function(X) t(apply(X, 1, function(x){x/sum(x)}))
prop.by.row <- function(X) sweep(X, 1, rowSums(X), "/")

ExZ.X <- function(p.vec, eta.mat, X){
  N <- nrow(X)
  C <- ncol(X)
  M <- length(p.vec)
# outer(X, 1:N, 1:M, function(x, i, j) p.vec[j] * dmultinom(X[i,], prob = eta.mat[j,]))
  Ez <- matrix(nrow = N, ncol = M)
  for(i in 1:N){
  for(j in 1:M){
    Ez[i, j] <- p.vec[j] * dmultinom(X[i,], prob = eta.mat[j,])
  }
  }
  prop.by.row(Ez)
}
```

Next, we construct function mparms. 
```{r}
Mparms <- function(expect.Z, X){
  eta.mat <- prop.by.row(t(expect.Z) %*% X)
  p.vec <- colMeans(expect.Z)
  list(eta = eta.mat, p = p.vec)
}
```

Following step 6 and 7,we test if there is any difference between the original value and the calculated one by trying out Exz.X as suggested. 
```{r test}
set.seed(121)
N <- 20
M <- 2
C <- 3
n <- 0.4*N
true.Z.vec <- as.vector(unlist(mapply(rep, c(1, 2), c(n, N-n))))
true.Z <- matrix(0, nrow = N, ncol = M)
true.eta <- round(prop.by.row(matrix(runif(M * C), nrow = M)), 2)
X <- matrix(nrow = N, ncol = C)
m.vec <- round(runif(N) * 100)
for(i in 1:N){
  true.Z[i, true.Z.vec[i]] <- 1
}
for(i in 1:N){
  X[i,] <- as.vector(rmultinom(1, size = m.vec[i], prob = true.Z[i,] %*% true.eta))
}

p.vec <- rep(1/M, M)
eta.mat <- prop.by.row(X)

for(it in 1:20){
expect.Z <- ExZ.X(p.vec, eta.mat, X)
expect.M <- Mparms(expect.Z, X)
p.vec <- expect.M$p
eta.mat <- expect.M$eta
}
# check ExZ.X
round(colMeans(expect.Z), 2)
# check eta.mat
round(expect.M$eta, 2)
true.eta
```
The Exz.X seems to return as expected. However, we still encounter some challenges to reach the best expected eta.mat.


# Contribution Statement
Both Alex and Kun wrote their own code of the function first and discuss on the final version.

Repository: [kuy003](https://github.com/Kun-Yang-003/FMPH291_WI22_A00_kuy003.git)
