---
title: "trimean"
author: "Kun Yang & Alexander Zhu"
date: "3/7/2022"
output: pdf_document
---

Conduct a simulation study to assess the bias and variance of the trimean for samples of size 9 from the
standard normal distribution and for t-distributions with 4 and 10 degrees of freedom.

Let us construct the Standard normal distribution as below: 
```{r}
#sample size is 9
n <- 9
# simulation 10000 times
N <- 10000

# mean of standard normal distribution 
mu_ <- 0
# standard deviation of standard normal distribution 
sd_ <- 1

set.seed(12)
#construct the trimean function that is an estimate of location based on the median and the quartiles that is constructed as a weighted average with the median weighted twice as heavily as each quartile.
Xtrimean <- rep(NA, N)
for(j in 1:N){
  Xn <- rnorm(n, mu_, sd_)
  Xn <- sort(Xn)
  Xtrimean[j] <- (Xn[3]+2*Xn[5]+Xn[7])/4
}

(E.Xtrimean <- mean(Xtrimean))
(E.Xtrimean.true <- (t(c(1, 2, 1)) %*% qnorm(c(1/3, 1/2, 2/3), mu_, sd_))[1])
(Var.Xtrimean <- var(Xtrimean))
# standard errors of estimates
(SE.Var.Xtrimean <- sqrt(var((Xtrimean-E.Xtrimean)^2)/N))
SE.Var.Xtrimean.1 <- sqrt(2/N)*Var.Xtrimean

```
Then we construct t-distribution with 4 degrees of freedom: 

```{r}
#sample size is 9
n <- 9
# simulation 10000 times
N <- 10000

# degree of freedom is 4
df_ <- 4

set.seed(12)
Xtrimean <- rep(NA, N)
for(j in 1:N){
  Xn <- rt(n, df_)
  Xn <- sort(Xn)
  Xtrimean[j] <- (Xn[3]+2*Xn[5]+Xn[7])/4
}

(E.Xtrimean <- mean(Xtrimean))
(E.Xtrimean.true <- (t(c(1, 2, 1)) %*% qt(c(1/3, 1/2, 2/3), df_))[1])
(Var.Xtrimean <- var(Xtrimean))
(SE.Var.Xtrimean <- sqrt(var((Xtrimean-E.Xtrimean)^2)/N))
SE.Var.Xtrimean.1 <- sqrt(2/N)*Var.Xtrimean
```

Next we construct a t-distribution with 10 degrees of freedom: 

```{r}
#sample size is 9
n <- 9
# simulation 10000 times
N <- 10000

# degree of freedom is 10
df_ <- 10

set.seed(12)
Xtrimean <- rep(NA, N)
for(j in 1:N){
  Xn <- rt(n, df_)
  Xn <- sort(Xn)
  Xtrimean[j] <- (Xn[3]+2*Xn[5]+Xn[7])/4
}

(E.Xtrimean <- mean(Xtrimean))
(E.Xtrimean.true <- (t(c(1, 2, 1)) %*% qt(c(1/3, 1/2, 2/3), df_))[1])
(Var.Xtrimean <- var(Xtrimean))
(SE.Var.Xtrimean <- sqrt(var((Xtrimean-E.Xtrimean)^2)/N))
SE.Var.Xtrimean.1 <- sqrt(2/N)*Var.Xtrimean
```
Based on the results, the standard error for the standard normal distribution is 0.001903596 which is the smallest compared to the t-distribution simulation results. For the t-distribution simulation, with the increase of degree of freedom from 4 to 10, the standard error decreases from  0.00261053 to 0.002079812. This is with our intuition that when the degree of freedom increases large enough, the simulation result of t-distribution would coincide with that of standard normal distribution. 


## Contribution Statement

Both Alex and Kun wrote their own code of the function first and discuss on the final version.

Repository: [kuy003](https://github.com/Kun-Yang-003/FMPH291_WI22_A00_kuy003.git)
