---
title: "truncsim"
author: "Kun Yang & Alexander ZHU"
date: "3/15/2022"
output: pdf_document
---

```{r}
library(randomTrunc)
```

The maximization procedure is implemented with `optim` function on the negative loglikelihoods with respect to the models.  

```{r}
N <- 1000
```

## Simulation 1
It is not easy to perform a grid search with 4 parameters, so we choose to generate 1000 sets of random parameters to evaluate the accuracy of the 2 models (all our conclusions will not change even if we change N to 10000).  

```{r}
n <- 20
mu_x <- 0.0
sigma_x <- 1.0
mu_z <- -1.0
sigma_z <- 0.10

set.seed(123)
x <- rtrunc4(n, mu_z, sigma_z, mu_x, sigma_x)

lst_parm <- list()
lst_parm$mu_x <- runif(N, -3, 3)
lst_parm$sigma_x <- runif(N, 0, 2)
lst_parm$mu_z <- runif(N, -2, 2)
lst_parm$sigma_z <- runif(N, 0, 1)
```

**Use 3 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 3)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "three", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_x"] - sigma_x)^2))
cat("\n")
```

**Use 4 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 4)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "four", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "sigma_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of sigma_z: ", mean((m_par[,"sigma_z"] - sigma_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_z"] - sigma_x)^2))
cat("\n")
```

For this simulation, the 3 parameters model works better than the 4 parameters model. It has higher accuracy and higher rate of convergence, and the program costs less time. But the 3 parameter method requires more appropriate choices on the initial values to avoid the failures at the initialization.  


## Simulation 2
 

```{r}
n <- 100
mu_x <- 0.0
sigma_x <- 1.0
mu_z <- -1.0
sigma_z <- 0.10

set.seed(123)
x <- rtrunc4(n, mu_z, sigma_z, mu_x, sigma_x)

lst_parm <- list()
lst_parm$mu_x <- runif(N, -3, 3)
lst_parm$sigma_x <- runif(N, 0, 2)
lst_parm$mu_z <- runif(N, -2, 2)
lst_parm$sigma_z <- runif(N, 0, 1)
```

**Use 3 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 3)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "three", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_x"] - sigma_x)^2))
cat("\n")
```

**Use 4 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 4)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "four", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "sigma_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of sigma_z: ", mean((m_par[,"sigma_z"] - sigma_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_z"] - sigma_x)^2))
cat("\n")
```

For this simulation, the 3 parameters model works better than the 4 parameters model. It has higher accuracy and higher rate of convergence, and the program costs less time. But the 3 parameter method requires more appropriate choices on the initial values to avoid the failures at the initialization.  


## Simulation 3
 

```{r}
n <- 1000
mu_x <- 0.0
sigma_x <- 1.0
mu_z <- -1.0
sigma_z <- 0.10

set.seed(123)
x <- rtrunc4(n, mu_z, sigma_z, mu_x, sigma_x)

lst_parm <- list()
lst_parm$mu_x <- runif(N, -3, 3)
lst_parm$sigma_x <- runif(N, 0, 2)
lst_parm$mu_z <- runif(N, -2, 2)
lst_parm$sigma_z <- runif(N, 0, 1)
```

**Use 3 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 3)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "three", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_x"] - sigma_x)^2))
cat("\n")
```

**Use 4 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 4)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "four", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "sigma_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of sigma_z: ", mean((m_par[,"sigma_z"] - sigma_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_z"] - sigma_x)^2))
cat("\n")
```

For this simulation, the 3 parameters model works better than the 4 parameters model. It has higher accuracy and higher rate of convergence, and the program costs less time. But the 3 parameter method requires more appropriate choices on the initial values to avoid the failures at the initialization.  
The errors of 4 parameters model get smaller when `n` is bigger.  

## Simulation 4
 

```{r}
n <- 1000
mu_x <- 0.0
sigma_x <- 1.0
mu_z <- -1.0
sigma_z <- 0.25

set.seed(123)
x <- rtrunc4(n, mu_z, sigma_z, mu_x, sigma_x)

lst_parm <- list()
lst_parm$mu_x <- runif(N, -3, 3)
lst_parm$sigma_x <- runif(N, 0, 2)
lst_parm$mu_z <- runif(N, -2, 2)
lst_parm$sigma_z <- runif(N, 0, 1)
```

**Use 3 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 3)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "three", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_x"] - sigma_x)^2))
cat("\n")
```

**Use 4 parameters:**  

```{r}
set.seed(123)
n_error <- 0
n_converge <- 0
m_par <- matrix(nrow = N, ncol = 4)
for(i in 1:N){
  out <- try(suppressWarnings(maxlik(x, model = "four", parm = c(lst_parm$mu_z[i], lst_parm$sigma_z[i], lst_parm$mu_x[i], lst_parm$sigma_x[i]))), silent = T)
  if(class(out) != "try-error"){
    if(out$convergence == 0){
      n_converge <- n_converge + 1
      m_par[i,] <- out$par
    }
  }else{
    n_error <- n_error + 1
  }
}
colnames(m_par) <- c("mu_z", "sigma_z", "mu_x", "sigma_x")
m_par <- na.omit(m_par)

cat("Number of failures due to inappropriate initial parameters: ", n_error, " (", round(n_error/N*100,2), "%)", sep = "")
cat("\n")
cat("Percent of convergence of the optim function: ", round(n_converge/(N-n_error)*100,2), "%", " (", n_converge, "/", N-n_error, ")", ", ", round(n_converge/(N)*100,2), "%", " (", n_converge, "/", N, ")", sep = "")
cat("\n")
cat("Only coverged results are used for calculating MSEs.")
cat("\n")

cat("MSE of mu_z: ", mean((m_par[,"mu_z"] - mu_z)^2))
cat("\n")
cat("MSE of sigma_z: ", mean((m_par[,"sigma_z"] - sigma_z)^2))
cat("\n")
cat("MSE of mu_x: ", mean((m_par[,"mu_x"] - mu_x)^2))
cat("\n")
cat("MSE of sigma_x: ", mean((m_par[,"sigma_z"] - sigma_x)^2))
cat("\n")
```


For this simulation, the 3 parameters model works better than the 4 parameters model. It has higher accuracy and higher rate of convergence, and the program costs less time. But the 3 parameter method requires more appropriate choices on the initial values to avoid the failures at the initialization.  

Although it is expected that if the true `sigma_z` gets bigger, the model with 4 parameters may have better accuracy than the 3 parameter model. According to the results saved in `m_par`, there is higher chance that the 4 parameter model converges to values that are far more different from the true value.  

