---
title: "lab3"
author: Kun Yang
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

# Floating Point Errors
```{r}
source("../R/fpSum.R")
```

## simulation on x and y
```{r}
set.seed(123)
x <- runif(50, min = 0, max = 1)
y <- runif(50, min = 0, max = 1)
```

```{r}
d <- 2
plot(density(abs(fpSum(x, y, digits = d) - (x + y))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(x, y, digits = d) - (x + y)))
```
The maximum value would be 0.05 if x and y were infinitely long.  

The maximum value would be closer to 0.05 if the vectors get longer.  
```{r, eval=FALSE}
set.seed(123)
x <- runif(1000, min = 0, max = 1)
y <- runif(1000, min = 0, max = 1)
# plot(density(abs(fpSum(x, y, digits = d) - (x + y))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(x, y, digits = d) - (x + y)))
```

### more digits
```{r}
d <- 3
plot(density(abs(fpSum(x, y, digits = d) - (x + y))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(x, y, digits = d) - (x + y)))
```
The maximum value would be 0.005 if x and y were infinitely long.  


```{r}
d <- 4
plot(density(abs(fpSum(x, y, digits = d) - (x + y))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(x, y, digits = d) - (x + y)))
```
The maximum value would be 0.0005 if x and y were infinitely long.  


### machine epsilon
$\beta$ corresponds to the floating point base, 10.  
$d$ corresponds to the number of precision, depending on the number of digits we use.  


### simulateion with z additionally
```{r}
z <- runif(50, min = 0, max = 1)
```

### digits = 2
```{r}
d <- 2
plot(density(abs(fpSum(fpSum(x, y, digits = d), z) - (x + y + z))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(fpSum(x, y, digits = d), z, digits = d) - (x + y + z)))
```
The maximum value would be 0.1 if x and y were infinitely long.  
So every time we apply the `fpSum` function, it will introduce an error with the maximum of 0.05 when `digits = 2`.  

### digits = 3
```{r}
d <- 3
plot(density(abs(fpSum(fpSum(x, y, digits = d), z) - (x + y + z))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(fpSum(x, y, digits = d), z, digits = d) - (x + y + z)))
```
The maximum value would be 0.01 if x and y were infinitely long.  

### digits = 4
```{r}
d <- 4
plot(density(abs(fpSum(fpSum(x, y, digits = d), z) - (x + y + z))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(fpSum(x, y, digits = d), z, digits = d) - (x + y + z)))
```

```{r}
set.seed(123)
x <- runif(1000, min = 0, max = 1)
y <- runif(1000, min = 0, max = 1)
z <- runif(1000, min = 0, max = 1)
d <- 4
plot(density(abs(fpSum(fpSum(x, y, digits = d), z) - (x + y + z))), main = paste0("digits = ", d))
# hist(abs(fpSum(x, y, digits = d) - (x + y)), main = paste0("digits = ", d))
max(abs(fpSum(fpSum(x, y, digits = d), z, digits = d) - (x + y + z)))
```
The maximum value would be 0.001 if x and y were infinitely long.  


# Better Finv*( … )

```{r}
source("../R/Finv.R")
```

When the value of `a` is greater, the precision of the value returned from `Finv0` gets worse.

```{r}
Finv1(0.5, 100)
```
`Finv1` doesn't return an expected outcome.


Using `log.p = TRUE` in `Finv2` to address this issue:
```{r}
Finv2 <- function(u, a) {
q <- pnorm(a, lower.tail = FALSE, log.p = T)
qnorm(q + log(1 - u), lower.tail = FALSE, log.p = T)
}
```

```{r}
all.equal(Finv1(u, 3), Finv2(u, 3))
```
`Finv2` returns exactly same values as `Finv1` when `a = 3`

```{r}
plot(u, Finv2(u, 100), type = "l")
```
