---
title: "longley"
author: "Kun Yang"
date: "1/27/2022"
output: pdf_document
---

## data
```{r}
# ? longley
longley.x <- data.matrix(longley[, 1:6])
longley.x.1 <- as.matrix(cbind(Intercept = 1, longley.x))
longley.y <- matrix(longley[, "Employed"], ncol = 1)
```


## part a. exact least squares

### a simple least squares fitting
```{r}
ls_fit <- function(x, y, output_type = "floating"){
  solve(t(x) %*% x) %*% t(x) %*% y
}
```

```{r}
ls_fit(longley.x.1, longley.y)
```

### introducing package `gmp`
```{r}
library(gmp)
ls_fit_bigq <- function(x, y, output_type = "floating"){
  x <- as.bigq(x)
  y <- as.bigq(y)
  output <- solve.bigq(t(x) %*% x) %*% t(x) %*% y
  if(output_type == "floating") output <- matrix(as.double(output), ncol = 1)
  if(output_type == "rational") output <- as.bigq(output)
  output
}
```

```{r}
round(ls_fit_bigq(longley.x.1, longley.y, output_type = "rational"), 2)
```

```{r}
(coef_ls_bigq <- ls_fit_bigq(longley.x.1, longley.y))
```


## part b. Cholesky factorization
```{r}
ls_chol <- function(x, y){
  L <- chol(crossprod(unname(x)))
  solve(L) %*% solve(t(L)) %*% t(x) %*% y
}
```

```{r}
ls_chol_bigq <- function(x, y, output_type = "floating"){
  L <- chol(crossprod(unname(x)))
  L <- as.bigq(L)
  x <- as.bigq(x)
  y <- as.bigq(y)
  output <- solve.bigq(L) %*% solve.bigq(t(L)) %*% t(x) %*% y
  if(output_type == "floating") output <- matrix(as.double(output), ncol = 1)
  if(output_type == "rational") output <- as.bigq(output)
  output
}
```


### without centering
```{r}
(coef_chol_bigq <- ls_chol_bigq(longley.x.1, longley.y))
```

### using centering
```{r}
longley.x.2 <- sweep(longley.x, 2, colMeans(longley.x))
longley.x.2 <- as.matrix(cbind(Intercept = 1, longley.x.2))
(coef_c_chol_bigq <- ls_chol_bigq(longley.x.2, longley.y))
```


## part c. comparison
```{r}
coef_lm <- coef(lm(Employed ~ ., data = longley))
coef_lm <- matrix(unname(coef_lm), ncol = 1)
```

To compare with the centered method, we exclude the intercept term in our comparison.

```{r}
mean((coef_ls_bigq - coef_lm)[-1,])
mean((coef_ls_bigq - coef_chol_bigq)[-1,])
mean((coef_ls_bigq - coef_c_chol_bigq)[-1,])
```

So, the non-centered Cholesky approach returns greatest biased estimation and centering will help to improve the accuracy.  

```{r, eval=FALSE, include=FALSE}
tolerance <- 1e-20
all.equal(coef_lm[-1,], coef_ls_bigq[-1,], tolerance = tolerance)
all.equal(coef_lm[-1,], coef_chol_bigq[-1,], tolerance = tolerance)
all.equal(coef_lm[-1,], coef_c_chol_bigq[-1,], tolerance = tolerance)
```

